### 2023-12-12 paper/spec.tex
Add spec file for the rules

\documentclass{article}

\usepackage{ottalt}
\usepackage{mathpartir}
\usepackage{supertabular}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}

\inputott{rules}

\title{System Specification}
\author{Yiyun Liu}

\begin{document}
\maketitle

\section{Grammar}
\ottgrammar

\section{Dynamics}
\label{sec:dynamics}
\drules[P]{$[[a => b]]$}{Parallel Reduction}{Var, Set, Void, Pi, Abs, App, AppAbs, True, False, If, IfTrue, IfFalse, Bool, Eq, Refl, J, JRefl}

\section{Statics}
\label{sec:statics}
\drules[T]{$[[G |- a : A]]$}{Typing}{Var, Set, Pi, Abs, App, Conv, Refl, J, Bool, True, False, If, Void}

\end{document}