### 2023-12-22 paper/spec.tex
Update spec file with semantic typing

\documentclass{article}

\usepackage{ottalt}
\usepackage{mathpartir}
\usepackage{supertabular}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}

\inputott{rules}

\title{System Specification}
\author{Yiyun Liu}

\begin{document}
\maketitle

\section{Grammar}
\ottgrammar

\section{Dynamics}
\label{sec:dynamics}
\drules[C]{$[[a <=> b]]$}{Coherence}{Intro}
\drules[PS]{$[[a =>+ b]]$}{Transitive Closure of Parallel Reduction}{One, Step}
\drules[P]{$[[a => b]]$}{Parallel Reduction}{Var, Set, Void, Pi, Abs, App, AppAbs, True, False, If, IfTrue, IfFalse, Bool, Eq, Refl, J, JRefl}

\section{Statics}
\label{sec:statics}
\drules[Ctx]{$[[|-G]]$}{Context Well-Formedness}{Empty, Cons}
\drules[T]{$[[G |- a : A]]$}{Typing}{Var, Set, Pi, Abs, App, Conv, Refl, J, Bool, True, False, If, Void}

\section{Semantic Typing}
\label{sec:semtyping}
\drules[I]{$[[I i A S]]$}{Logical Relation}{Void, Bool, Eq, Pi, Set, Red}

\end{document}